--- 
title: "Прикладной анализ данных"
author: "Елена Рыбина"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography:
- book.bib
- packages.bib
description: |
  Занятие по прикладному анализу данных.
link-citations: yes
github-repo: elenary/AppAnDan
---
```{r, eval=TRUE,echo=FALSE}
library(tidyverse)
library(viridis)
```
# Чек-лист для работы с данными

1. Сформулированы теоретические гипотезы
2. Сформулированы эмпирические гипотезы
3. Есть понимание, как будут выглядеть / выглядят собранные данные
4. Есть понимание, что данных достаточно для проверки поставленной гипотезы
5. Исходя из гипотезы и того, как собраны и какие данные, выбран статистический метод проверки экспериментальной гипотезы
6. Исходя из статистического метода, есть понимание о статистической гипотезе *(хорошо, если есть)*
7. Выбрана среда для анализа данных и пакет для этого метода и конкретная функция
8. Данные приведены в вид, к которому можно применить выбранный статистический метод
9. Визуализация самих данных для лучшего их понимания
10. Проведен расчет выбранного стат метода
11. Вид полученных результатов понятен для интерпретации, и резульатты согласуются с представлением о том, как они должны выглядеть
12. Результаты проинтепретированы и стедлан вывод относительно изначальной гипотезы
13. Визуализированы результаты, так, чтобы их было легче правильно проинтерпретировать


## Сформулированы теоретические гипотезы 

Возьмем датасет с kaggle <https://www.kaggle.com/datasets/uciml/student-alcohol-consumption>
```{r, eval=TRUE}
studens1 <- read_csv("student-mat.csv")
View(studens1)
```

```{r, eval=TRUE}
studens1 %>% 
  rename_with(., ~ paste0(., "_mat"), .cols = c(absences, paid, G1, G2, G3)) -> studens1
```

```{r, eval=TRUE}
studens2 <- read_csv("student-por.csv")
studens2 %>% 
  rename_with(., ~ paste0(., "_por"), .cols = c(absences, paid, G1, G2, G3)) -> studens2
View(studens2)
```


Each **bookdown** chapter is an .Rmd file, and each .Rmd file can contain one (and only one) chapter. A chapter *must*
The `index.Rmd` file is required, and is also your first book chapter. It will be the homepage when you render the book.


## Render book

You can render the HTML version of this example book without changing anything:

1. Find the **Build** pane in the RStudio IDE, and

1. Click on **Build Book**, then select your output format, or select "All formats" if you'd like to use multiple formats from the same book source files.

Or build the book from the R console:

```{r, eval=FALSE}
bookdown::render_book()
```

To render this example to PDF as a `bookdown::pdf_book`, you'll need to install XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.org/tinytex/>.

## Preview book

As you work, you may start a local server to live preview this HTML book. This preview will update as you edit the book when you save individual .Rmd files. You can start the server in a work session by using the RStudio add-in "Preview book", or from the R console:

```{r eval=FALSE}
bookdown::serve_book()
```


```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
